### Data Preparation

#### Step1: Data Acquisition

```{r}
library(readr)
X2019 <- read_csv("/Users/kachunlee/Downloads/2019.csv")
View(X2019)
```

#### Step2: Explore data

***1. summary***

```{r}

summary(X2019)
```


#### Step3: Deal with missing values:

```{r}
sapply(X2019, function(x) sum(is.na (X2019)))
```

There are no na value. good.

#### Step4: ouliers


We use boxplot to drop the points below Q1 - 1.5 \* IQR or above Q3 + 1.5 \* IQR.

```{r}
bx1<-boxplot(X2019$`GDP per capita`, ylab = "Healthy life expectancy")
```

no outliers for this column.

```{r}
bx2<-boxplot(X2019$`Social support`, ylab = "Social support")
```

There are some outliers

```{r}
out_ind2 <- which( X2019$`Social support`%in% bx2$out)
X2019_no<-X2019[-out_ind2,]
dim(X2019_no)
```

```{r}
bx3<-boxplot(X2019_no$`Healthy life expectancy`, ylab = "Health (Life Expectancy)")

```

1 outliers!

```{r}
out_ind3 <- which(X2019_no$`Healthy life expectancy`%in% bx3$out)
X2019_no<-X2019_no[-out_ind3,]
dim(X2019_no)
```

```{r}
dim(X2019)
```


```{r}
bx4<-boxplot(X2019_no$`Healthy life expectancy`, ylab = "Healthy life expectancy")
```



```{r}
bx5<-boxplot(X2019_no$`Freedom to make life choices`, ylab = "Freedom")
```


```{r}
bx6<-boxplot(X2019_no$`Freedom to make life choices`, ylab = "Freedom")
```


```{r}
out_ind6 <- which( X2019_no$`Freedom to make life choices`%in% bx6$out)
X2019_no<-X2019_no[-out_ind6,]
dim(X2019_no)
```

```{r}
bx7<-boxplot(X2019_no$Generosity, ylab = "Generosity")
```


```{r}
out_ind7 <- which( X2019_no$Generosity%in% bx7$out)
X2019_no<-X2019_no[-out_ind7,]
dim(X2019_no)
```

delete all the outliers now!!!

```{r}
bx8<-boxplot(X2019_no$`Perceptions of corruption`, ylab = "Corruption")
```

```{r}
out_ind8 <- which( X2019_no$`Perceptions of corruption`%in% bx8$out)
X2019_no<-X2019_no[-out_ind8,]
dim(X2019_no)
```

#### Step5: correlation/collinearity/distribution analysis

Use pairs.panels to see the distributions and correlation of each feature in the data set.

```{r}
library(psych)
pairs.panels(X2019_no,method = "pearson")
```

#### Step6: normalization

```{r}
#X2019_scaled<-X2019_no
#X2019_scaled
#use scale function to normalize all the columns except the target
#X2019_scaled<-X2019_scaled[, colnames(X2019_scaled)[c(2:8)]]
#X2019_scaled$Happyiness<-X2019_no$`Happiness Score`
#X2019_scaled

```

check out if the columns have been normalized

```{r}
#summary(X2015_scaled)
```

```{r}
#lin<-lm(Happyiness~. , data = X2015_scaled)
#summary(lin)
```

```{r}
library(readr)
X2020 <- read_csv("/Users/kachunlee/Downloads/2020.csv")
View(X2020)
```

#### Step2: Explore data

***1. summary***

```{r}

summary(X2020)
```

#### Step3: Deal with missing values:

```{r}
sapply(X2020, function(x) sum(is.na (X2020)))
```

There are no na value. good.

#### Step4: ouliers

We use boxplot to drop the points below Q1 - 1.5 \* IQR or above Q3 + 1.5 \* IQR.

```{r}
bx1<-boxplot(X2020$`Ladder score`, ylab = "Happiness-Score")
```

no outliers for column 1.

```{r}
bx2<-boxplot(X2020$`Logged GDP per capita`, ylab = "Logged GDP per capita")
```

no outliers for column 2.

```{r}
bx3<-boxplot(X2020$'Social support', ylab = "Family")
```

2 outliers!

```{r}
out_ind3 <- which( X2020$'Social support'%in% bx3$out)
X2020_no<-X2020[-out_ind3,]
dim(X2020_no)
```

```{r}
dim(X2020)
```

ok, delete the outliers for coolumn3!

```{r}
bx4<-boxplot(X2020_no$`Healthy life expectancy`, ylab = "Health (Life Expectancy)")
```

no outliers!

```{r}
bx5<-boxplot(X2020_no$'Freedom to make life choices', ylab = "Freedom")
```

2 outliers!

out_ind2 <- which( X2020$'Freedom to make life choices'%in% bx5$out)
X2020_no<-X2020_no[-out_ind2,]
dim(X2020_no)

```{r}
bx6<-boxplot(X2020_no$`Generosity`, ylab = "Generosity")
```

2 outliers!

```{r}
out_ind6 <- which( X2020_no$`Generosity`%in% bx6$out)
X2020_no<-X2020_no[-out_ind6,]
dim(X2020_no)
```

```{r}
bx7<-boxplot(X2020_no$'Perceptions of corruption', ylab = "Perceptions of corruption")
```

many outlier!

```{r}
out_ind7 <- which( X2020_no$'Perceptions of corruption'%in% bx7$out)
X2020_no<-X2020_no[-out_ind7,]
dim(X2020_no)
```

delete all the outliers now!!!

#### Step5: correlation/collinearity/distribution analysis

Use pairs.panels to see the distributions and correlation of each feature in the data set.

```{r}
library(psych)
pairs.panels(X2020_no,method = "pearson")
```

```{r}
dim(X2019_no)
dim(X2020_no)
```

```{r}
mean(X2019_no$Score)
mean(X2020_no$`Ladder score`)
```


```{r}
wilcox.test(X2019_no$Score, X2020_no$`Ladder score`, paired = TRUE, alternative ="two.sided")
```
```{r}
Y2019 <- read_csv("/Users/kachunlee/Downloads/2019_1.csv")
Y2020 <- read_csv("/Users/kachunlee/Downloads/2020_1.csv")
Y<-read_csv("/Users/kachunlee/Downloads/combine.csv")
```
```{r}
wilcox.test(Y2019$Score, Y2020$`Ladder score`, paired = TRUE, alternative ="two.sided")
```

```{r}
t.test(X2019$Score,X2020$`Ladder score`,paired = FALSE)
```




```{r}
Y2019
Y2020
Y
```

```{r}
install.packages(ggpubr)
library(ggpubr)
```


```{r}
ggboxplot(data = Y, x = 'group', y = 'Score', 
       palette = c("#00AFBB", "#E7B800"),
          ylab = "score", xlab = "year")
```

```{r}
wilcox.test(Y2019$Score, Y2020$`Ladder score`, alternative = "two.sided")
```

```{r}
df <- read_csv("/Users/kachunlee/Downloads/anova3.csv")
df
```

```{r}
model <- aov(Score~group, data=df)
summary(model)
```

```{r}
TukeyHSD(model, conf.level=.95)
plot(TukeyHSD(model, conf.level=.95), las = 2)
```